\section{Second-order design}\label{sec:second-order}

With armature inductance \(L_a\), armature resistance \(R_a\), inertia \(J\),
viscous friction \(b\), and motor constant \(K_m\), the continuous-time
second-order transfer function is
\begin{equation}
  G(s)=
  \frac{K_m}
       {L_aJs^2+\left(L_ab+R_aJ\right)s+\left(R_ab+K_m^2\right)} .
  \label{eq:second-order-continuous-transfer}
\end{equation}
After zero-order-hold discretisation this is written as
\begin{equation}
  G(z)=\frac{b_1z+b_0}{z^2+a_1z+a_0}.
  \label{eq:second-order-discrete-transfer}
\end{equation}
The adaptive estimator therefore has four parameters, with regression vector
\begin{equation}
  \varphi(k)=
  \begin{bmatrix}
    -y(k-1) & -y(k-2) & u(k-1) & u(k-2)
  \end{bmatrix}.
  \label{eq:second-order-regressor}
\end{equation}

The numerical values used here,
\[
  K_m=0.05,\quad
  J=\SI{0.001}{kg.m^2},\quad
  b=0.0005,\quad
  L_a=\SI{2}{mH},\quad
  R_a=\SI{2}{\ohm},
\]
are representative parameters of a small brushed DC motor, not measurements of
the 25GA371 rig motor. The conclusion does not depend on their exact values,
only on the electrical time constant being far shorter than the sampling time.
They give continuous poles \(-998.7\,\si{s^{-1}}\) and
\(-1.75\,\si{s^{-1}}\), so the electrical time constant is about
\(\SI{1}{ms}\). Table~\ref{tab:second-order-zoh-values} shows the exact ZOH
discretisation at the hardware sampling time and at a sampling time that
resolves the electrical pole.

\begin{scripttable}
  \centering
  \caption{Exact ZOH parameters of the second-order motor model.}
  \label{tab:second-order-zoh-values}
  \begin{tabular}{rrrrrrrl}
    \toprule
    \(T\) & \(a_1\) & \(a_0\) & \(b_1\) & \(b_0\) &
    \(b_0/b_1\) & \multicolumn{2}{c}{\(z\)-poles} \\
    \midrule
    \(\SI{80}{ms}\) & \(-0.8692\) & \(0.000000\) & \(1.8467\) &
    \(0.0218\) & \(\SI{1.18}{\percent}\) & \(0.8692\) & \(0\) \\
    \(\SI{2}{ms}\)  & \(-1.1322\) & \(0.1352\)   & \(0.0284\) &
    \(0.0148\) & \(\SI{52.3}{\percent}\) & \(0.9965\) & \(0.1357\) \\
    \bottomrule
  \end{tabular}
\end{scripttable}

Example 7.6 of the reader does not use the free \(c(z)\) of the general
Diophantine system; it forces \(c(z)=z-1\), i.e. a PI controller. The
second-order design does the same thing one order higher. Pure pole placement
fixes only the poles, not the static gain, and left a
\(\SI{17.2}{\percent}\) steady-state error in simulation. To remove this error,
the design forces
\begin{equation}
  c(z)=(z-1)(z+c_0),
  \label{eq:second-order-integrator}
\end{equation}
which gives \(T(1)=1\). Equation 7.16 of the reader gives the design equation
whose linear system is solved for the controller coefficients at every
sampling instant.

The desired closed-loop polynomial is
\begin{equation}
  q(z)=z(z-0.1357)(z-0.9608)^2 .
  \label{eq:second-order-desired-polynomial}
\end{equation}
The fast electrical pole is deliberately left in place in
\eqref{eq:second-order-desired-polynomial} because it is already much faster
than the desired bandwidth. If instead
\[
  q(z)=(z-0.9608)^4
\]
is demanded, the design slows the electrical pole by a factor of \(50\). The
controller must then invert the fast plant dynamics and becomes unstable
itself, with its pole at \(|z|=2.66\), and behind a saturation limit such a
controller necessarily diverges. The design rule is therefore to leave modes
that are much faster than the desired bandwidth where they are.

The working design uses \(T=\SI{2}{ms}\), with PRBS dither added to the
manipulated variable. Clipping does not corrupt the estimate because the
regressor uses the post-saturation command. The estimator is used without
forgetting, \(\alpha=1\), because without persistent excitation \(\alpha<1\)
makes \(P\) grow without bound.

\subsection{Identifiability on the rig}\label{sec:identifiability}

The ARX model is structurally identifiable at both sampling times. At
\(T=\SI{80}{ms}\), \(A(z)=z(z-0.8692)\) and
\(B(z)=1.8467z+0.0218\) are coprime; the ZOH numerator zero is at
\(-b_0/b_1=-0.0118\), so it does not cancel the pole at the origin. The
failure on the rig is practical, or statistical, identifiability.

Two effects pull against each other with \(T\). Sensitivity to the electrical
mode decays as \(\exp(p_eT)\), so at \(T=\SI{80}{ms}\) the \(b_0\) term is only
\(\SI{1.18}{\percent}\) of the output, while the speed quantisation step
\(d_q=2\pi/(T\,\mathrm{CPR})\) grows as \(1/T\), with variance proportional to
\(1/T^2\). With the measured \(\mathrm{CPR}=44\), no sampling time satisfies
both requirements.

\begin{scripttable}
  \centering
  \caption{Open-loop second-order identification under encoder quantisation.}
  \label{tab:second-order-identifiability-study}
  \begin{tabular}{lrrrrr}
    \toprule
    setting & step [\si{rad/s}] & \(a_1\) err & \(b_1\) err &
    \(\lvert a_0\ \mathrm{err}\rvert\) & \(b_0\) err \\
    \midrule
    \multicolumn{6}{l}{Encoder resolution at \(T=\SI{80}{ms}\)} \\
    \(\mathrm{CPR}=44\)  & \(1.785\) & \(-25.6\,\%\) & \(-0.06\,\%\) & \(0.194\) & \(1882\,\%\) \\
    \(\mathrm{CPR}=704\) & \(0.112\) & \(-0.12\,\%\) & \(0.00\,\%\)  & \(0.0009\) & \(9.1\,\%\) \\
    exact                & --        & \(0.00\,\%\)  & \(0.00\,\%\)  & \(1.4\times10^{-5}\) & \(0.13\,\%\) \\
    \midrule
    \multicolumn{6}{l}{Sampling time at \(\mathrm{CPR}=44\)} \\
    \(T=\SI{80}{ms}\) & \(1.785\)  & \(-25.6\,\%\)  & \(-0.06\,\%\)  & \(0.194\) & \(1882\,\%\) \\
    \(T=\SI{20}{ms}\) & \(7.140\)  & \(-49.8\,\%\)  & \(-0.97\,\%\)  & \(0.462\) & \(934\,\%\) \\
    \(T=\SI{5}{ms}\)  & \(28.560\) & \(-43.5\,\%\)  & \(-11.2\,\%\)  & \(0.382\) & \(129\,\%\) \\
    \(T=\SI{2}{ms}\)  & \(71.400\) & \(-100.0\,\%\) & \(-100.0\,\%\) & \(0.135\) & \(-100\,\%\) \\
    \bottomrule
  \end{tabular}
\end{scripttable}

Table~\ref{tab:second-order-identifiability-study} uses open-loop, full-order
PRBS data precisely to remove closed-loop bias and poor excitation as
explanations, leaving the sensor. The \(a_0\) column is an absolute error
because the true value at \(T=\SI{80}{ms}\) is numerically zero, so a relative
error would be meaningless. Quantisation also corrupts the regressors
\(y(k-1)\) and \(y(k-2)\) in \eqref{eq:second-order-regressor}, an
errors-in-variables problem that biases the estimate rather than merely adding
variance. This is why the hardware estimates settle on wrong
values instead of scattering about the right ones.

At \(T=\SI{2}{ms}\), \(\mathrm{CPR}=44\) gives only about \(1.3\) encoder
counts per sample at a moderate \(\SI{90}{rad/s}\), and \(10.9\) counts per
sample at the measured full-duty \(\SI{780}{rad/s}\), so the speed signal
degenerates towards a one-bit signal. This is pure arithmetic from two
measured constants and holds regardless of the motor model.

The mechanical pair is robust: \(b_1\) is still recovered to
\(\SI{0.06}{\percent}\) at \(\mathrm{CPR}=44\) and \(T=\SI{80}{ms}\). That is
why the first-order design of Example 7.6 runs on this hardware and the
second-order extension does not. The controller only needs the
well-determined combinations of the parameters, not their individual values,
so the closed loop stays stable. A stable closed loop is therefore not evidence of successful identification.
